---
title: "Part 2: Adding Multiple Diseases & Tools"
---

`epiworldR` supports multi-variant models, the below code gives instructions on 
how to implement this. First, build a `SIRCONN` model for COVID-19, which new
viruses and tools will then be added to. 

## Adding Multiple Viruses & Variants

```{r design-and-add}
library(epiworldR)
model_sir <- ModelSIRCONN(
  name              = "COVID-19",
  n                 = 100000, 
  prevalence        = 0.0001, 
  contact_rate      = 2,
  prob_transmission = 0.5,
  prob_recovery     = 1/3
  )
run(model_sir, ndays = 50, seed = 1912)
model_sir
```

#### Designing the Virus

Using the `virus()` function, assign a name the new virus/variant with its corresponding probability of infecting any given agent. In this example, prob_infecting = 0.3. In order to add this new virus to the model, use the `add_virus()` function by calling the original `epiworldR` model object, the new virus, and the new virus' prevalence (which is set to 0.0001 in this example).

```{r}
# Building the virus
flu <- virus(name = "Flu", prob_infecting = .3)

# Adding the virus to the model
add_virus(model_sir, flu, .0001)
```

Next, run the updated model with the new virus for 50 days, the output below describes the simulation. To confirm that the flu is included, notice the presence of "Flu" in the Virus(es) section of the output. All other output is interpretable as specified in previous sections.

```{r}
run(model_sir, ndays = 50, seed = 1912)
model_sir
```

#### Plotting
Plotting the previous model (including the flu) yields the following. Notice the presence of two reproductive numbers plotted over time. Variant 0 refers to COVID-19 and variant 1 refers to the flu.

```{r, fig.height=10}
repnum2 <- get_reproductive_number(model_sir)

op <- par(mfrow = c(2,1))
plot(model_sir)
plot(repnum2, type="b")
par(op)
```

## Using tools

Now, the implementation of tools to fight any viruses and variants in the model will be demonstrated. First, for simplicity, remove the flu virus from the `SIR` model object (keep in mind the index for the flu virus in the model object is 1). Next, provide parameters for the new tool using the `tool()` function. These parameters include the tool's name, any reduction in probabilities for the `SIR` model parameters, and increased probability of recovery option. To add the tool to the `SIR` model, use the `add_tool()` function with the `SIR` model object, new tool, and prevalence of the tool. In this example, assume that 50% of the population will have received the vaccination.

```{r removing-virus}
# Removing the flu virus from the model
rm_virus(model_sir, 1)

vaccine <- tool(
  name = "Vaccine",
  susceptibility_reduction = .9,
  transmission_reduction = .5,
  recovery_enhancer = .5, 
  death_reduction = .9
)

add_tool(model_sir, vaccine, 0.5)
run(model_sir, ndays = 50, seed = 1231)
```

#### Plotting

```{r curves-including-vaccine, fig.height=10}
repnum3 <- get_reproductive_number(model_sir)

op <- par(mfrow = c(2,1))
plot(model_sir)
plot(repnum3, type="b")
par(op)
```
## Comorbidities Using Logit Functions

Many times we want to model the effects of comorbidities on the
disease. For example, we may want to model the effects of obesity on the
probability of recovery from COVID-19. To do this, we can use the `virus_fun_logit()` function
to model the probability of recovery. 

The steps are the following:

1. Create the model

2. Assign the agents' data (a matrix with covariates/features) to the model.

3. Create a function to model the probability of recovery using the `virus_fun_logit()` function.

4. Add the function to the virus' recovery rate using `set_prob_recovery_fun()`.

5. Run the model.

We start by creating two matching models, one with comorbidities and one without.

```{r}
# With comorbidities
model_comor <- ModelSEIRCONN(
  name              = "Flu",
  n                 = 10000, 
  prevalence        = 0.001, 
  contact_rate      = 2.1,
  prob_transmission = 0.5,
  incubation_days   = 7,
  prob_recovery     = 1/4
  )

# Without comorbidities
model_no_comor <- ModelSEIRCONN(
  name              = "Flu",
  n                 = 10000, 
  prevalence        = 0.001, 
  contact_rate      = 2.1,
  prob_transmission = 0.5,
  incubation_days   = 7,
  prob_recovery     = 1/4
  )
```

Next, we will create a matrix with the agents' data. As an example, we will
create a matrix with two columns, one for the baseline and one for obesity.

```{r}
# Artificial population with obesity
set.seed(88)
X <- cbind(
  baseline = 1,
  obese    = sample.int(2, size(model_comor), replace = TRUE) - 1
  )

# Looking at the data
head(X)
```

Let's now link agents' data to the model. This will allow us to use the data to
model the probability of recovery.

```{r}
# Adding the data to the model
set_agents_data(model_comor, X)
model_comor
```

We then use `virus_fun_logit()` to create a function we can use to model the
probability of recovery. The function takes in the following arguments:

- `vars`: the variables to use in the model
- `coefs`: the coefficients for each variable
- `model`: the model object

```{r}
# Logit function 
lfun <- virus_fun_logit(
  vars  = 0:1,
  coefs = c(
    qlogis(.25),
    qlogis(.25 - .25/2)
    ), 
  model_comor
  )

# Printing
lfun
```

The next step is to set the probability of recovery function for the virus. We
can do this using the `set_prob_recovery_fun()` function:

```{r}
# Setting the probability of recovery
set_prob_recovery_fun(
  virus = get_virus(model_comor, 0), 
  model = model_comor,
  vfun  = lfun
  )
```

We are now ready to run the model.

```{r}
run(model_comor, ndays = 50, seed = 1231)
run(model_no_comor, ndays = 50, seed = 1231)
```

And see the result

```{r}
#| label: curves-comorbidities
op <- par(mfrow = c(1, 2), cex = .7)
plot_incidence(model_comor, main = "With comorbidities")
plot_incidence(model_no_comor, main = "Without comorbidities")
par(op)
```

Let's also look at the states of the agents in the model using the `get_state()` and
`get_agents()` functions.

```{r}
states_comor <- sapply(1:size(model_comor), \(i) {
  get_state(get_agents(model_comor)[i - 1])
})

states_no_comor <- sapply(1:size(model_comor), \(i) {
  get_state(get_agents(model_no_comor)[i - 1])
})

table(states_comor, states_no_comor)

# We can even compute the correlation
t.test(X[,2],
  (states_no_comor < 3) - 
  (states_comor < 3)
  )
```

The table above shows that the model with comorbidities has fewer
recovered agents than the model without comorbidities. Furthermore, the model
with no comorbidities still has susceptible agents, which is not the case
for the model with comorbidities.

## Exercise 1

Using a `SIRCONN` model to simulate the Flu for 75 days, add the Coronavirus 
Delta variant to the model. Then plot the model parameters and reproductive 
numbers over time for both viruses. Assume n = 10000, prevalence = 0.0001, 
contact_rate = 2.1, prob_transmission = 0.5, and prob_recovery = $\frac{1}{4}$ 
for model initialization. Assume the Delta variant has a prob_infecting = 0.3,
prob_recovery = $\frac{1}{4}$, and an initial prevalence = 0.001.  
When running the model, use seed = 1912.
  
After how many days does the number of infections peak in this simulation? 
How many infections occur at the peak?

:::{.callout-note collapse="true"}
## Solution
```{r, fig.height=10}
model_sir <- ModelSIRCONN(
  name              = "Flu",
  n                 = 10000, 
  prevalence        = 0.0001, 
  contact_rate      = 2.1,
  prob_transmission = 0.5,
  prob_recovery     = 1/4
  )

# Building the virus
delta <- virus(name = "Delta", prob_infecting = .3, prob_recovery = 1/4)

# Adding the virus to the model
add_virus(model_sir, delta, .001)

# Run the model
run(model_sir, ndays = 75, seed = 1912)

# Reproductive number 
repnum <- get_reproductive_number(model_sir)

# Plotting
op <- par(mfrow = c(2,1))
plot(model_sir)
plot(repnum, type="b")
par(op)

# Day of peak infections 
x <- get_hist_total(model_sir)
which.max(x$counts[x$state == "Infected"]) - 1
# Number of infections on peak day
max(x$counts[x$state == "Infected"])
```
:::

## Exercise 2

Using the `SIRCONN` model from exercise 1 (model_sir), remove the Delta variant
from the model and add a masking tool called "mask_wearing" with transmission 
reduction = 0.3 and proportion of complying agents = 0.6. Simulate this 
scenario for 50 days and plot the model parameters and average reproductive number
over time. When running the model, use seed = 1912.

After how many days does the number of infections peak in this simulation?  
How many infections occur at the peak?

:::{.callout-tip collapse="true"}
Masking only influences the transmission of a disease, thus transmission reduction 
= 0.3, and all other parameters of this tool will be 0.0. 
:::

:::{.callout-note collapse="true"}
## Solution
```{r, fig.height=10}
# Removing Delta variant
rm_virus(model_sir, 1)

# Creating a tool
mask_wearing <- tool(
  name = "Mask",
  susceptibility_reduction = 0.0,
  transmission_reduction   = 0.3, # Only transmission
  recovery_enhancer        = 0.0,
  death_reduction          = 0.0
)

# Adding tool to the model
add_tool(model_sir, mask_wearing, 0.6)

# Running the model
run(model_sir, ndays = 50, seed = 1912)

# Reproductive number
repnum2 <- get_reproductive_number(model_sir)

# Plotting
op <- par(mfrow = c(2,1))
plot(model_sir)
plot(repnum2, type="b")
par(op)

# Day of peak infections
y <- get_hist_total(model_sir)
which.max(y$counts[y$state == "Infected"]) - 1
# Number of infections on peak day
max(y$counts[y$state == "Infected"])
```
:::
