---
title: "Part 2a: Social Network Analysis"
---

## Network Diffusion

epiworldR also supports diffusion networks to simulate the spread of ideas or rumors throughout a population, for example. The below example demonstrates this capability.

#### Example

Goal: Create a diffusion network model which simulates the spread of a rumor throughout the UKfaculty population of size n = 81.

```{r, include = FALSE}
library(igraph)
library(netplot)
```

```{r}
# First, load in the dataset using the igraph library
# library(igraph)
data(UKfaculty, package="igraphdata")

# Visualizing the Network
# library(netplot)
nplot(UKfaculty)
```

The above plot visualizes the UK faculty network of size 81 and its interaction patterns.

```{r}
set.seed(2223)

# Turning the network into an edgelist
UKfaculty_vertex_data <- as_data_frame(UKfaculty, what = c("vertices"))
UKfaculty_matrix <- as.matrix(UKfaculty_vertex_data)
UKfaculty_edgelist <- as_edgelist(UKfaculty)

# Simulate a 2 column numeric matrix and use for data in ModelDiffNet
# Creating the diffusion model
library(epiworldR)
adopt_rumor <- ModelDiffNet(
  "Rumor",
  prevalence = .08,
  prob_adopt = .1,
  data       = UKfaculty_matrix,
  params     = c(1, 4)
)

n <- igraph::vcount(UKfaculty)
g <- matrix(runif(n ^ 2) < .01, nrow = n)
diag(g) <- FALSE
el <- which(g, arr.ind = TRUE) - 1L

# Simulating a population from smallworld - Use agents_from_edgelist
agents_from_edgelist(adopt_rumor,
                     source = el[,1],
                     target = el[,2],
                     size = n, 
                     directed = TRUE)

# Running the model for 50 steps
run(adopt_rumor, 50)

# Plotting the model
plot(adopt_rumor)
```

The number of rumor-adopters steadily increases over time as the number of rumor non-adopters decreases, an expected result when considering a social network.
