---
title: "Part 2a: Non-Disease Applications - Network Diffusion"
---

Though `epiworldR` focuses primarily on disease modeling, the `epiworld` framework is general enough to model any type of diffusion process, such as simulating the spread of ideas or rumors throughout a population. In this part of the workshop, we'll walk through an example of social network diffusion.

## Example Scenario: Academic Faculty Rumor

The example implements the following scenario:

-   **Rumor:** The dean will resign!
-   **Population Description:** A UK university faculty
-   **Population size:** 81 agents

### Load Packages and Data

We use the `igraph` and `netplot` R packages for manipulating and plotting networks respectively. The file `part2a.rda` includes the `UKfaculty` friendship network from the `igraphdata` package, with the vertex attributes (`UKfaculty_vertex`) and edge list (`UKfaculty_edgelist`) as numeric matrices. We load the `UKfaculty` data and visualize the friendship network.

```{r include=FALSE}
library(igraph)  # For manipulating networks
library(netplot) # For plotting
```

```{r}
load("part2a.rda")

nplot(UKfaculty)
```

### Simulating the Rumor

Using the `ModelDiffNet()` function, we can simulate how the rumor spreads through the network. First, we create the network diffusion model. The prevalence of the rumor is set such that 2 agents are exposed to the rumor at the start of the simulation. The probability of an agent adopting the rumor is 0.1.

```{r}
library(epiworldR)

adopt_rumor <- ModelDiffNet(
  name       = "The dean will resign!",
  prevalence = 2/vcount(UKfaculty),
  prob_adopt = .1,
  data       = UKfaculty_vertex,
  params     = c(0, 5)
)
```

We then initialize the model agents using the `agents_from_edgelist()` function.

```{r}
# Reading in the network
agents_from_edgelist(
  adopt_rumor,
  size     = vcount(UKfaculty),
  source   = UKfaculty_edgelist[,1] - 1L,
  target   = UKfaculty_edgelist[,2] - 1L,
  directed = TRUE
  )
```

We run the model for 100 days and plot the results.

```{r}
verbose_off(adopt_rumor)
run(adopt_rumor, ndays = 100)

adopt_rumor

plot(adopt_rumor)
```

By day 16 or 17, the entire network had adopted the rumor that the dean will resign.

### View the Adoption Network

The transmission network is not always relevant for a diffusion network model, but in this case we can use it to visualize the spread of the rumor through the network. We get the transmission network using the `get_transmissions()` function, then plot the network using the `igraph` and `netplot` packages.

```{r}
transmissions <- get_transmissions(adopt_rumor)

# Turning it into a network via igraph
tnet <- graph_from_data_frame(
  transmissions[, c(2, 3)] + 1, # To start from 1 (not 0)
  directed = TRUE,
  vertices = cbind(id = 1:81, UKfaculty_vertex)
  )

# Adding back vertex attributes
nplot(tnet)
```
