---
title: "Part 3: Multiple Simulation Runs"
---

```{r, echo = FALSE}
# Set the seed for reproducibility
set.seed(123)

# Define the range of the uniform distribution
lower <- 0
upper <- 6

# Initialize vectors to store the sample sizes and their corresponding averages
sample_sizes <- c()
averages <- c()

# Set the maximum sample size
max_n <- 1000

# Generate samples and calculate averages for increasing sample sizes
for (n in 1:max_n) {
  # Generate a random sample from the uniform distribution
  sample <- runif(n, min = lower, max = upper)
  # Calculate the average of the sample
  average <- mean(sample)
  # Store the sample size and the corresponding average
  sample_sizes <- c(sample_sizes, n)
  averages <- c(averages, average)
}
plot(sample_sizes, averages, type = "l", xlab = "Sample Size (n)", 
     ylab = "Average", main = "Average vs. Sample Size")
```

In statistics, the Law of Large Numbers ensures that as the sample size increases, the sample mean (average) of a random variable will converge to the population mean. The same principle applies when using multiple runs to simulate epidemiological models. As the number of `epiworldR` simulations increases, the sample means of the reproductive number or model parameters, for example, will converge to their corresponding population means.

In `epiworldR`, the `run_multiple()` function allows us to run a specified number of simulations using the same model object to compare model results across several separate and repeated simulations. In this part of the workshop, we'll walk through an example of using this function.

## Example Scenario: 500 Simulations of Measles

The example implements the following scenario:

-   **Disease:** Measles
-   **Population size:** 5,000 agents
-   **Initial Disease Prevalence:** 0.001
-   **Contact Rate:** 2
-   **Transmission Rate:** 0.5
-   **Incubation Days:** 7
-   **Recovery Rate:** $\frac{1}{7}$

We'll run 500 simulations of this model and compare the results.

### Model Setup

Use the `ModelSEIRCONN()` function to initialize the measles model.

```{r}
#| label: model-setup
library(epiworldR)

measles_model <- ModelSEIRCONN(
  name              = "Measles",
  n                 = 5000, 
  prevalence        = 0.001, 
  contact_rate      = 2,
  transmission_rate = 0.5,
  incubation_days   = 7,
  recovery_rate     = 1/7
  )
```

### Generate Saver

Use the `make_saver()` function to create a saver object. The saver will extract the `total_hist` and `reproductive` data from the model object after each run so the results can be compared.

::: callout-note
You can generate a saver for any metric compatible with the `make_saver()` function. See the details section of the `make_saver()` help manual for more (`?make_saver`).
:::

```{r}
#| label: generate-saver
measles_saver <- make_saver("total_hist", "reproductive")
```

### Run 500 Simulations

Use the `run_multiple()` function to simulate the measles model 500 times (`nsims = 500`) with each simulation running for 50 days (`ndays = 50`). Add the saver created in the previous step and set the number of threads to 4 (`nthread = 4`) for parallel computing.

```{r}
#| label: run-model-multiple
verbose_off(measles_model)
run_multiple(measles_model, ndays = 50, nsims = 500, saver = measles_saver, nthread = 4)
```

Note that we do not run the model with a random seed, because we want each model run to have random differences.

### Visualize Results

Create a boxplot of the four SEIR states for each day across the 500 simulations using the `run_multiple_get_results()` function. For this example, we filter the data to observe only the first 20 days.

```{r}
#| label: plot-epicurv-multiple
seirconn_500 <- run_multiple_get_results(measles_model)$total_hist
seirconn_500 <- seirconn_500[seirconn_500$date <= 20,]
plot(seirconn_500)
```

Create a boxplot of the reproductive number for all 50 days across the 500 simulations using the `run_multiple_get_results()` function.

```{r}
#| label: plot-reproductive-number
seirconn_500_r <- run_multiple_get_results(measles_model)$reproductive
plot(seirconn_500_r)
```

As expected, the reproductive number on average, decreases over time.

## Exercise 1

Consider for this exercise that there is a Hepatitis A outbreak. Your goal is to observe the average reproductive number over 100 simulations. Using a `run_multiple` simulation, what is the average reproductive number over the course of the first 20 days? Use a `SEIRCONN` model with:\
- n = 10000\
- prevalence = 0.01\
- contact_rate = 2\
- transmission_rate = 0.5\
- incubation_days = 2\
- recovery_rate = $\frac{1}{7}$

::: {.callout-tip collapse="true"}
General Steps:\
1. Create `epiworldR_model`\
2. Generate saver\
3. Run `epiworldR_model`\
4. Plot average reproductive number
:::

```{r}
# Your solution here

```

## Exercise 2

Using the same `SEIRCONN` model from exercise 1, simulate a vaccine intervention for the previous exercise's Hepatitis A virus outbreak where 50% of individuals in the population will receive the vaccine on day 10. How then, does the average reproductive number behave over 20 days and 100 simulations? Assume the following parameters:\
- susceptibility_reduction = .9\
- transmission_reduction = .5\
- recovery_enhancer = .5\
- death_reduction = .9

::: {.callout-tip collapse="true"}
General Steps:\
1. Create tool\
2. Use `globalaction_tool` & `add_global_action`\
3. Generate saver & `run_multiple`\
4. Plot average reproductive number
:::

```{r}
# Your solution here

```
